Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-14 preliminaries.tex
Add notes and others' updates

Let $\mathcal{B}$ be the Hilbert space corresponding to one qubit. Let $H:\mathcal{B}^{\otimes n}\rightarrow\mathcal{B}^{\otimes n}$ be Hermitian matrices. We use $H\geq0$ to denote $H$ being positive semidefinite. Let $\lambda(H)$ be the smallest eigenvalue of $H$. The ground states of $H$ are the eigenvectors corresponding to $\lambda(H)$. For matrix $H$ and subspace $S$, let $H\big|_S=\Pi_S H \Pi_S$, where $\Pi_S$ is the projector onto the subspace $S$. For a $T$-qubit Hilbert space, let the state $\ket{\widehat{t}}=\ket{1}^{\otimes t}\otimes \ket{0}^{{\otimes (T-t)}}$.
For $n$-qubit states $\rho_1, \rho_2\in\cB^{\otimes n}$,
we write $F(\rho_1, \rho_2)=\left(\tr\sqrt{\sqrt{\rho_1}\rho_2\sqrt{\rho_1}}\right)^2$ for the fidelity between $\rho_1$ and $\rho_2$.
We write $\norm{\rho_1-\rho_2}_{tr}=\frac{1}{2}\norm{\rho_1-\rho_2}_1$ for the trace distance between $\rho_1$ and $\rho_2$,
We have $\norm{\rho_1-\rho_2}_{tr}\leq\sqrt{1-F(\rho_1, \rho_2)}$.
We say $\rho_1$ and $\rho_2$ are $\varepsilon$-close if $F(\rho_1, \rho_2)\leq\varepsilon$.
For two distributions $f_1$ and $f_2$ over the same finite domain $X$, we write $\norm{f_1-f_2}_{TV}=\frac{1}{2}\sum_{x\in X}\abs{f_1(x)-f_2(x)}$ for their total variation distance,
and we say $f_1$ and $f_2$ are $\varepsilon$-close if $\norm{f_1-f_2}_{TV}\leq\varepsilon$.
We write $x\xleftarrow{\$}X$ when $x$ is sampled uniformly at random from the set $X$.

@@

\begin{definition} [$\SampBQP$]
	$\SampBQP$ is the class of sampling problems $\left(D_x\right)_{x\in\set{0, 1}^*}$ that can be (approximately) sampled by polynomial-size uniform quantum circuits. Namely, there exists a Turing machine $M$ such that for every $n \in \bbN$ and $\varepsilon \in (0,1)$, $M(1^n, 1^{1/\eps})$ outputs a quantum circuit $C$ in $\poly(n, 1/\varepsilon)$ time such that for every $x \in \zo^n$, the output of $C(x)$ (measured in standard basis) is $\eps$-close to $D_x$.
\end{definition}

@@

@@

\begin{definition}
	\label{dfn:stats-secure-proto-sampbqp}
	Let $\Pi=(P, V)$ be a $\QPIP_\tau$ protocol.
	We say it is a protocol for the $\SampBQP$ instance $(D_x)_{x\in\zo^*}$ with completeness error $c$ and statistical (resp., computational) soundness if the following holds:
	\begin{itemize}
		\item On public inputs $1^\lambda$, $1^{1/\varepsilon}$, and $x\in\zo^n$, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$. If $d=\Acc$ then $z\in\zo^{m(n)}$, otherwise $z=\bot$.
		\item (Completeness):
			For all security parameter $\lambda\in\bbN$, accuracy parameter $\varepsilon=\frac{1}{\poly(\lambda)}$, and $x\in\zo^{\poly(\lambda)}$, 
			Let $(d, z)\leftarrow(P, V)(1^\lambda, 1^{1/\varepsilon}, x)$, then $d=\Rej$ with probability at most $c(\lambda)$.
		\item (Statistical soundness): For all cheating prover $P^*$,
			sufficiently large $\lambda\in\bbN$, accuracy parameter $\varepsilon=\frac{1}{\poly(\lambda)}$, and $x\in\zo^{\poly(\lambda)}$, 
			consider the following experiment:
			\begin{itemize}
				\item Let $(d, z)\leftarrow(P^*, V)(x, 1^{1/\varepsilon}, 1^\lambda)$.
				\item Define $z_{ideal}$ by
				$$\begin{cases}
					z_{ideal}=\bot & \text{if } d=\Rej\\
					z_{ideal}\leftarrow D_x & \text{if } d=\Acc
				\end{cases}$$.
			\end{itemize}
			It holds that $\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\varepsilon$. 
		\item (Computational soundness): Define $d, z, z_{ideal}$ same as above.
			It holds that for all $\BQP$ distinguisher $D$,
			$\abs{\Pr[D(1^{\lambda}, d,z) = 1] - \Pr[D(1^{\lambda}, d,z_{ideal}) = 1]} \leq \varepsilon $
	\end{itemize}
\end{definition}

We now define blindness of a protocol.
In the context of blind protocols, the above definitions should be modified so that $x$ is the verifier's private input.
We define blindness with an accuracy parameter $\varepsilon$, which exists in protocols for $\SampBQP$ instances;
one can leave it out where appropriate, such as in protocols for $\BQP$ languages.

\begin{definition}
	Let $\Pi=(P, V)$ be an interactive protocol with common inputs $1^\lambda$ and $1^{1/\varepsilon}$, and verifier's private input $x\in\zo^*$.
	We say $\Pi$ is \emph{blind} if for all cheating prover $P^*$ the following ensembles are computationally indistinguishable over $\lambda$:
	\begin{itemize}
		\item $\set{\View_{P^*}(P^*, V(x))(1^\lambda, 1^{1/\varepsilon})}_{\lambda\in\bbN, \varepsilon\in(0,1), x\in\zo^*}$
		\item $\set{\View_{P^*}(P^*, V(0))(1^\lambda, 1^{1/\varepsilon})}_{\lambda\in\bbN, \varepsilon\in(0,1), x\in\zo^*}$
	\end{itemize}
\end{definition}

@@

\Ethan{Proj. lemma and de Finetti thm. were moved}

@@

We use the quantum fully homomorphic encryption scheme given in \cite{mahadev_qfhe} which is compatible with our use of a classical client.
\begin{definition}
	A quantum homomophic (public-key) encryption scheme $\QHE=(\QGen,$ \linebreak
	$\QEnc, \QEval, \QDec)$ is quadruple of quantum polynomial-time algorithms which operate as follows:
	\begin{itemize}
		\item \emph{Key generation.}
			The algorithm $(pk, sk)\leftarrow\QGen(1^\lambda)$ takes a unary representation of the security parameter and outputs a public key encryption key $pk$ and a secret decryption key $sk$.
		\item \emph{Encryption.}
			The algorithm $c\leftarrow\QEnc(pk, \mu)$ takes the public key $pk$ and a single bit message $\mu\in\zo$ and outputs a ciphertext $c$.
		\item \emph{Decryption.}
			The algorithm $\mu^*\leftarrow\QDec(sk, c)$ takes the secret key $sk$ and a ciphertext $c$ and outputs a message $\mu^*\in\zo$.
			It must be the case that
				$$\QDec(sk, \QEnc(pk, \mu))=\mu$$
			with overwhelming probability in $\lambda$.
		\item \emph{Homomorphic Evaluation.}
			The algorithm $c_f\leftarrow\QEval(pk, f, c_1, \ldots, c_l)$ takes the public key $pk$, a quantum circuit \Ethan{?} $f$, and a set of $l$ ciphertexts $c_1,\cdots,c_l$ and outputs a ciphertext $c_f$.
			It must be the case that
				$$\QDec(sk, \QEval(pk, f, c_1, \ldots, c_l))=f(\QDec(sk, c_1),\ldots,\QDec(sk, c_l))$$
			with overwhelming probability in $\lambda$.
	\end{itemize}
\end{definition}

\begin{definition}[Compactness and Full Homomorphism]
	\Ethan{This is copied from Mahadev. Not sure how much of this we need. We need at least how leveled HE is defined.}

	A quantum homomorphic encryption scheme $\QHE$ is compact if
	there exists a polynomial $s$ in $\lambda$ such that the output length of $\QEval$ is at most $s$ bits long (regardless of $f$ or
	the number of inputs). A compact scheme is (pure) fully homomorphic if it can evaluate any efficiently computable
	boolean function. A compact scheme is leveled fully homomorphic if it takes $1^L$ as additional input in key generation,
	and can only evaluate depth $L$ Boolean circuits where $L$ is polynomial in $\lambda$. A scheme is quantum pure or
	leveled fully homomorphic if we refer to quantum circuits rather than boolean circuits.
\end{definition}

@@

\begin{definition}
	\Ethan{This was copied from Mahadev. Not sure where $L$ (level) goes.}
	A scheme $\QHE$ is IND-CPA secure if, for any polynomial time adversary $\cA$, there exists a negligible function $\mu(\cdot)$ such that
	$$\abs{Pr[\cA(pk, \mathsf{HE.Enc}_{pk}(0))=1]-Pr[\cA(pk, \mathsf{HE.Enc}_{pk}(1))=1]}\leq\mu(\lambda)$$
	where $(pk, sk)\leftarrow\QGen(1^\lambda)$
\end{definition}

@@

\begin{definition}
	we call a quantum homomorphic encryption scheme \emph{classical-friendly} if it has the following properties:
	\begin{itemize}
		\item $\QGen$ produces classical keys
		\item In the case where the plaintext is classical, $\QEnc$ can be done classically.
		\item When the underlying message is classical, the corresponding ciphertext is also classical and can be decrypted classically.
	\end{itemize}
\end{definition}

\begin{theorem} [Theorem 1.1 in \cite{mahadev_qfhe}]
	Under the assumption that the learning with errors problem with superpolynomial noise ratio is computationally intractable for an efficient quantum machine,
	there exists a classical-friendly quantum homomorphic encryption scheme.
\end{theorem}

>>>>

Let $\mathcal{B}$ be the Hilbert space corresponding to one qubit. Let $H:\mathcal{B}^{\otimes n}\rightarrow\mathcal{B}^{\otimes n}$ be Hermitian matrices. We use $H\geq0$ to denote $H$ being positive semidefinite. Let $\lambda(H)$ be the smallest eigenvalue of $H$. The ground states of $H$ are the eigenvectors corresponding to $\lambda(H)$. For matrix $H$ and subspace $S$, let $H\big|_S=\Pi_S H \Pi_S$, where $\Pi_S$ is the projector onto the subspace $S$. For a $T$-qubit Hilbert space, let the state $\ket{\widehat{t}}=\ket{1}^{\otimes t}\otimes \ket{0}^{{\otimes (T-t)}}$.
For $n$-qubit states $\rho_1, \rho_2\in\cB^{\otimes n}$,
we write $F(\rho_1, \rho_2)=\left(\tr\sqrt{\sqrt{\rho_1}\rho_2\sqrt{\rho_1}}\right)^2$ for the fidelity between $\rho_1$ and $\rho_2$.
We write $\norm{\rho_1-\rho_2}_{tr}=\frac{1}{2}\norm{\rho_1-\rho_2}_1$ for the trace distance between $\rho_1$ and $\rho_2$,
We have $\norm{\rho_1-\rho_2}_{tr}\leq\sqrt{1-F(\rho_1, \rho_2)}$.
We say $\rho_1$ and $\rho_2$ are $\varepsilon$-close if $\norm{\rho_1- \rho_2}_{tr}\leq\varepsilon$.
For two distributions $f_1$ and $f_2$ over the same finite domain $X$, we write $\norm{f_1-f_2}_{TV}=\frac{1}{2}\sum_{x\in X}\abs{f_1(x)-f_2(x)}$ for their total variation distance,
and we say $f_1$ and $f_2$ are $\varepsilon$-close if $\norm{f_1-f_2}_{TV}\leq\varepsilon$.
We write $x\xleftarrow{\$}X$ when $x$ is sampled uniformly at random from the set $X$.

@@

\begin{definition} [$\SampBQP$]
	$\SampBQP$ is the class of sampling problems $\left(D_x\right)_{x\in\set{0, 1}^*}$ that can be (approximately) sampled by polynomial-size uniform quantum circuits. Namely, there exists a Turing machine $M$ such that for every $n \in \bbN$ and $\varepsilon \in (0,1)$, $M(1^n, 1^{1/\varepsilon})$ outputs a quantum circuit $C$ in $\poly(n, 1/\varepsilon)$ time such that for every $x \in \zo^n$, the output of $C(x)$ (measured in standard basis) is $\varepsilon$-close to $D_x$.
\end{definition}

@@

\Ethan{Use Mahadev's definitions, but add eps param

change to subnormalized states

Need to define what it means to apply algorithm to subnormalized states

Two families of (potentially sub-normalized) states... leave author note for applying algorithms to subnormalized states; might be ok

change ``attacker" to ``distinguisher"

sec. param change from n to lambda}

\Ethan{Comment out entire section including triangle ineq.; just leave what's said above}

@@

\begin{definition}
	\label{dfn:stats-secure-proto-sampbqp}
	Let $\Pi=(P, V)$ be a $\QPIP_\tau$ protocol.
	We say it is a protocol for the $\SampBQP$ instance $(D_x)_{x\in\zo^*}$ with completeness error $c(\cdot)$ and statistical (resp., computational) soundness if the following holds:
	\begin{itemize}
		\item On public inputs $1^\lambda$, $1^{1/\varepsilon}$, and $x\in\zo^{\poly(\lambda)}$, $V$ outputs $(d, z)$ where $d\in\set{\Acc, \Rej}$. If $d=\Acc$ then $z\in\zo^{m(\abs{x})}$, \footnote{$m$ is the length of... TODO} otherwise $z=\bot$.
		\item (Completeness):
			For all accuracy parameter $\varepsilon(\lambda)=\frac{1}{\poly(\lambda)}$,
			security parameter $\lambda\in\bbN$,
			and $x\in\zo^{\poly(\lambda)}$, 
			Let $(d, z)\leftarrow(P, V)(1^\lambda, 1^{1/\varepsilon}, x)$, then $d=\Rej$ with probability at most $c(\lambda)$.
		\item (Statistical soundness): For all cheating prover $P^*$,
		    accuracy parameter $\varepsilon(\lambda)=\frac{1}{\poly(\lambda)}$,
			sufficiently large $\lambda\in\bbN$, and $x\in\zo^{\poly(\lambda)}$, 
			consider the following experiment:
			\begin{itemize}
				\item Let $(d, z)\leftarrow(P^*, V)(1^\lambda, 1^{1/\varepsilon}, x)$.
				\item Define $z_{ideal}$ by
				$$\begin{cases}
					z_{ideal}=\bot & \text{if } d=\Rej\\
					z_{ideal}\leftarrow D_x & \text{if } d=\Acc
				\end{cases}$$.
			\end{itemize}
			It holds that $\norm{(d,z)-(d,z_{ideal})}_{TV}\leq\varepsilon$. 
		\item (Computational soundness):
		For all cheating $\BQP$ prover $P^*$, $\BQP$ distinguisher $\mathsf{D}$, accuracy parameter $\varepsilon(\lambda)=\frac{1}{\poly(\lambda)}$,
			sufficiently large $\lambda\in\bbN$, and all $x\in\zo^{\poly(\lambda)}$, 
			define $d, z, z_{ideal}$ by the same experiment as above.
			It holds that
			$$\abs{\Pr[\mathsf{D}(1^{\lambda}, d,z) = 1] - \Pr[\mathsf{D}(1^{\lambda}, d,z_{ideal}) = 1]} \leq \varepsilon $$
	\end{itemize}
\end{definition}

We now define blindness of a protocol.
In the context of blind protocols, we consider $x$ as the verifier's private input instead of a common input.  
%the above definitions should be modified so that $x$ is the verifier's private input.
%We define blindness with an accuracy parameter $\varepsilon$, which exists in protocols for $\SampBQP$ instances;
%one can leave it out where appropriate, such as in protocols for $\BQP$ languages.

\begin{definition}
	Let $\Pi=(P, V)$ be an interactive protocol with common inputs $1^\lambda$ (and $1^{1/\varepsilon}$ for the case of $\SampBQP$ problem), and verifier's private input $x\in\zo^*$.
	We say $\Pi$ is \emph{blind} if for all cheating $\BQP$ prover $P^*$ the following ensembles are computationally indistinguishable over $\lambda$:
	\begin{itemize}
		\item $\set{\View_{P^*}(P^*, V(x))(1^\lambda, 1^{1/\varepsilon})}_{\lambda\in\bbN, \varepsilon\in(0,1), x\in\zo^*}$
		\item $\set{\View_{P^*}(P^*, V(0))(1^\lambda, 1^{1/\varepsilon})}_{\lambda\in\bbN, \varepsilon\in(0,1), x\in\zo^*}$
	\end{itemize}
\end{definition}

@@

@@

We use the quantum fully homomorphic encryption scheme given in \cite{mahadev_qfhe} which is compatible with our use of a classical client.
\begin{definition}
	A quantum homomophic (public-key) encryption scheme $\QHE=(\QGen,$ \linebreak
	$\QEnc, \QEval, \QDec)$ is quadruple of quantum polynomial-time algorithms which operate as follows:
	\begin{itemize}
		\item \emph{Key generation.}
			The algorithm $(pk, sk)\leftarrow\QGen(1^\lambda)$ takes a unary representation of the security parameter as input and outputs a public key $pk$ and a secret key $sk$.
		\item \emph{Encryption.}
			The algorithm $c\leftarrow\QEnc(pk, \mu)$ takes the public key $pk$ and a single bit message $\mu\in\zo$ as inputs and outputs a ciphertext $c$.
		\item \emph{Decryption.}
			The algorithm $\mu^*\leftarrow\QDec(sk, c)$ takes the secret key $sk$ and a ciphertext $c$ as inputs and outputs a message $\mu^*\in\zo$.
			It must be the case that
				$$\QDec(sk, \QEnc(pk, \mu))=\mu$$
			with overwhelming probability in $\lambda$.
		\item \emph{Homomorphic Evaluation.}
			The algorithm $c_f\leftarrow\QEval(pk, f, c_1, \ldots, c_l)$ takes the public key $pk$, a quantum circuit $f$, and a set of $l$ ciphertexts $c_1,\cdots,c_l$ as inputs and outputs a ciphertext $c_f$.
			It must be the case that
				$$\QDec(sk, \QEval(pk, f, c_1, \ldots, c_l))=f(\QDec(sk, c_1),\ldots,\QDec(sk, c_l))$$
			with overwhelming probability in $\lambda$.
	\end{itemize}
\end{definition}

\begin{definition}[Compactness and Full Homomorphism]
	A quantum homomorphic encryption scheme $\QHE$ is compact if
	there exists a polynomial $s$ in $\lambda$ such that the output length of $\QEval$ is at most $s$ bits long (regardless of $f$ or
	the number of inputs). A compact scheme is (pure) fully homomorphic if it can evaluate any efficiently computable
	boolean function. A compact scheme is leveled fully homomorphic if it takes $1^L$ as additional input in key generation,
	and can only evaluate depth $L$ Boolean circuits where $L$ is polynomial in $\lambda$.
\end{definition}

\Ethan{Combine leveled dfn. with QFHE; keep compactness here}

@@

\begin{definition}
	A scheme $\QHE$ is IND-CPA secure if, for any polynomial time adversary $\cA$, there exists a negligible function $\mu(\cdot)$ such that
	$$\abs{Pr[\cA(pk, \mathsf{HE.Enc}_{pk}(0))=1]-Pr[\cA(pk, \mathsf{HE.Enc}_{pk}(1))=1]}\leq\mu(\lambda)$$
	where $(pk, sk)\leftarrow\QGen(1^\lambda)$
\end{definition}

@@

\begin{definition}
	we call a quantum homomorphic encryption scheme \emph{classical-friendly} if it has the following properties:
	\begin{itemize}
		\item $\QGen$ is a classical algorithm
		\item In the case where the plaintext is classical, $\QEnc$ can be done classically.
		\item When the underlying message is classical, the corresponding ciphertext is also classical and can be decrypted classically.
	\end{itemize}
\end{definition}

\begin{theorem} [Theorem 1.1 in \cite{mahadev_qfhe}]
	Under the assumption that the learning with errors problem with superpolynomial noise ratio is computationally intractable for an efficient quantum machine,
	there exists a classical-friendly compact quantum leveled homomorphic encryption scheme.
\end{theorem}

### 2020-04-14 qpip.tex
Add notes and others' updates

\section{Delegation Protocol for $\QPIP_1$ Client}

In this section, we construct a one-message $\QPIP_1$ delegation protocol for $\SampBQP$.
At a high level, it is a cut-and-choose protocol.
The server constructs multiple copies of the ground state as certificates,
then the client randomly chooses a copy to output and checks the rest.
Unfortunately our approach incurs inverse polynomial soundness errors.

There are two main challenges to this cut-and-choose approach.
First, the client needs to reliably extract the circuit output from its corresponding ground state.
We accomplish this by padding the circuit with identity gates at the end.
By doing so, the clock register collapses to after the last non-identity gate with high probability.

The other challenge is that a cheating prover's certificates can be arbitrarily entangled,
so common techniques such as Chernoff bounds can't be applied as-is.
This wasn't a challenge for $\Piblind$ for $\BQP$ earlier because...
\Ethan{TODO; need brief and intuitive explanation}
In our case with $\SampBQP$, we overcome this challenge instead by using de Finetti's theorem
to approximate our measurement results with that of some unentangled copies,
within inverse polynomial errors.

@@

	Ingrediants:
	\begin{itemize}
		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$. \Ethan{Refer to the appropriate previous theorem}
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $s=n+T+\frac{6T}{\varepsilon}$ be the number of qubits in $\ket{\psi_{C'(x)}}$
		\item Let $\cVGS$ be the verification algorithm for $H_{C'(x)}$ as defined in \Cref{AlgGroundStateCheck}.
	\end{itemize}

>>>>

%\section{Delegation Protocol for $\QPIP_1$ Client}

% In this section, we construct a one-message $\QPIP_1$ delegation protocol for $\SampBQP$.
% At a high level, it is a cut-and-choose protocol.
% The server constructs multiple copies of the ground state as certificates,
% then the client randomly chooses a copy to output and checks the rest.
% Unfortunately our approach incurs inverse polynomial soundness errors.

% There are two main challenges to this cut-and-choose approach.
% First, the client needs to reliably extract the circuit output from its corresponding ground state.
% We accomplish this by padding the circuit with identity gates at the end.
% By doing so, the clock register collapses to after the last non-identity gate with high probability.

% The other challenge is that a cheating prover's certificates can be arbitrarily entangled,
% so common techniques such as Chernoff bounds can't be applied as-is.
% This wasn't a challenge for $\Piblind$ for $\BQP$ earlier because...
% \Ethan{TODO; need brief and intuitive explanation}
% In our case with $\SampBQP$, we overcome this challenge instead by using de Finetti's theorem
% to approximate our measurement results with that of some unentangled copies,
% within inverse polynomial errors.

@@

	Ingredients:
	\begin{itemize}
		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$. \Ethan{Refer to the appropriate previous theorem}
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $s=n+T+\frac{6T}{\varepsilon}$ be the number of qubits in $\ket{\psi_{C'(x)}}$
		\item Let $\cVGS$ be the verification algorithm for $H_{C'(x)}$ as defined in \Cref{AlgGroundStateCheck}.
	\end{itemize}
